Illustrating imputation of missing data using the Titanic dataset

### 2024-07-07 text/23/too-many-colors.rmd
Updated 2024 Rmd files

```{r, echo=FALSE, warning=FALSE}
library(usmap)
library(ggplot2)
plot_usmap(
    data = statepov, 
    values = "pct_pov_2014", 
    color = "red") -> us
us +
  scale_fill_continuous(
        low = "white", 
        high = "red", 
        name = "Poverty rate (2014)", 
        label = scales::comma) + 
  theme(legend.position = "right")
```

@@

```{r, echo=FALSE, warning=FALSE}
gr <- c("green", "yellow", "red")
us +
  scale_fill_gradientn(
        colors=gr, 
        name = "Poverty rate (2014)", 
        label = scales::comma) + 
  theme(legend.position = "right")
```

@@

```{r, echo=FALSE, warning=FALSE}
gr <- c("red", "yellow", "green", "cyan", "blue", "magenta")
us +
  scale_fill_gradientn(
        colors=gr, 
        name = "Poverty rate (2014)", 
        label = scales::comma) + 
  theme(legend.position = "right")
```

>>>>

```{r too-many-colors-01, echo=FALSE, warning=FALSE}
library(usmap)
library(ggplot2)
plot_usmap(
    data = statepov, 
    values = "pct_pov_2014", 
    color = "red") -> us
us +
  scale_fill_continuous(
        low = "white", 
        high = "red", 
        name = "Poverty rate (2014)", 
        label = scales::comma) + 
  theme(legend.position = "right")
```

@@

```{r too-many-colors-02, echo=FALSE, warning=FALSE}
gr <- c("green", "yellow", "red")
us +
  scale_fill_gradientn(
        colors=gr, 
        name = "Poverty rate (2014)", 
        label = scales::comma) + 
  theme(legend.position = "right")
```

@@

```{r too-many-colors-03, echo=FALSE, warning=FALSE}
gr <- c("red", "yellow", "green", "cyan", "blue", "magenta")
us +
  scale_fill_gradientn(
        colors=gr, 
        name = "Poverty rate (2014)", 
        label = scales::comma) + 
  theme(legend.position = "right")
```

### 2024-07-07 text/23/what-is-a-partial-correlation.Rmd
Updated 2024 Rmd files

---
title: "What is a partial correlation?"
source: "New"
author: Steve Simon
date: 2023-12-05
categories:
- Blog post
tags:
- Correlation
- Definitions
output: html_document
---

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(broom)
library(tidyverse)
fn <- "https://raw.githubusercontent.com/pmean/classes/master/biostats-1/data/data-05-fev.csv"
read_csv(fn, col_names=FALSE) %>%
  data.frame -> init
names(init) <- c(
  "age",
  "fev",
  "height",
  "sex",
  "smoker")
```

@@

```{r, echo=FALSE}
rxy <- round(cor(init$fev, init$age), 2)
rxz <- round(cor(init$fev, init$height), 2)
ryz <- round(cor(init$age, init$height), 2)
```

@@

```{r, echo=FALSE}
ggplot(init, aes(age, fev, label=height)) +
  geom_text() +
  ggtitle(paste0("Correlation = ", rxy))
h <- 48
```

@@

```{r, echo=FALSE}
g <- 4*floor((init$height-2)/4)+4
control_height <- function(a) {
  rxy <- round(cor(init$age[g==h], init$fev[g==h]), 2)
  co=ifelse(g==h, "black", "gray")
  al=ifelse(g==h, 1, 0)
  init %>%
    mutate(co=ifelse(g==h, "black", "gray")) %>%
    ggplot(aes(age, fev, label=height)) +
      geom_text(color=co) +
      geom_text(color=co, alpha=al) +
      ggtitle(paste0("Correlation = ", rxy))
}
h <- 48
control_height(h)
h <- h+4
```

@@

```{r, echo=FALSE}
control_height(h)
h <- h+4
```

@@

```{r, echo=FALSE}
control_height(h)
h <- h+4
```

@@

```{r, echo=FALSE}
control_height(h)
h <- h+4
```

@@

```{r, echo=FALSE}
control_height(h)
h <- h+4
```

@@

```{r, echo=FALSE}
control_height(h)
h <- h+4
```

@@

```{r, echo=FALSE}
control_height(h)
h <- h+4
```

>>>>

---
title: "What is a partial correlation?"
source: "New"
author: Steve Simon
date: 2023-12-05
categories:
- Blog post
tags:
- Correlation
- Definitions
- Rmarkdown graphs
output: html_document
page_update: complete
---

@@

```{r partial-correlation-01, echo=FALSE, warning=FALSE, message=FALSE}
library(broom)
library(tidyverse)
fn <- "https://raw.githubusercontent.com/pmean/classes/master/biostats-1/data/data-05-fev.csv"
read_csv(
    fn, 
    col_names=FALSE, 
    col_types="n") %>%
    data.frame -> init
names(init) <- c(
  "age",
  "fev",
  "height",
  "sex",
  "smoker")
```

@@

```{r partial-correlation-02, echo=FALSE}
rxy <- round(cor(init$fev, init$age), 2)
rxz <- round(cor(init$fev, init$height), 2)
ryz <- round(cor(init$age, init$height), 2)
```

@@

```{r partial-correlation-03, echo=FALSE}
ggplot(init, aes(age, fev, label=height)) +
  geom_text() +
  ggtitle(paste0("Correlation = ", rxy))
h <- 48
```

@@

```{r partial-correlation-04, echo=FALSE}
g <- 4*floor((init$height-2)/4)+4
control_height <- function(a) {
  rxy <- round(cor(init$age[g==h], init$fev[g==h]), 2)
  co=ifelse(g==h, "black", "gray")
  al=ifelse(g==h, 1, 0)
  init %>%
    mutate(co=ifelse(g==h, "black", "gray")) %>%
    ggplot(aes(age, fev, label=height)) +
      geom_text(color=co) +
      geom_text(color=co, alpha=al) +
      ggtitle(paste0("Correlation = ", rxy))
}
h <- 48
control_height(h)
h <- h+4
```

@@

```{r partial-correlation-05, echo=FALSE}
control_height(h)
h <- h+4
```

@@

```{r partial-correlation-06, echo=FALSE}
control_height(h)
h <- h+4
```

@@

```{r partial-correlation-07, echo=FALSE}
control_height(h)
h <- h+4
```

@@

```{r partial-correlation-08, echo=FALSE}
control_height(h)
h <- h+4
```

@@

```{r partial-correlation-09, echo=FALSE}
control_height(h)
h <- h+4
```

@@

```{r partial-correlation-10, echo=FALSE}
control_height(h)
h <- h+4
```

### 2024-07-07 text/24/diamond-spline.Rmd
Updated 2024 Rmd files

```{r initialization, echo=FALSE}
library(broom)
library(glue)
library(pspline)
library(readr)
library(rms)
library(tidyverse)
library(yaml)

@@

```{r read-diamonds}
f2 <- "singapore-diamond-prices.txt"
raw_data <- read.fwf(
  file=glue("{f0}/{f1}/{f2}"), 
  widths=c(5,2,5,4,5))
di <- raw_data
names(di) <- c(
  "carat",
  "colour",
  "clarity",
  "certification",
  "price"
)
```

@@

```{r adjustments}
exchange_rate <- 1.72
inflation_rate <- 1.81
di$adjusted_price <- di$price*inflation_rate/exchange_rate
head(di)
```

@@

```{r descriptives}
summary(di$carat)
summary(di$price)
summary(di$adjusted_price)
count(di, colour)
count(di, clarity)
count(di, certification)
```

@@

```{r plot}
ggplot(data=di, aes(carat, adjusted_price)) +
  geom_point(pch=1) -> scatter1
scatter1
```

@@

```{r linear-trend}
scatter1+geom_smooth(method="lm")
```

@@

```{r linear-fit}
linear_fit <- lm(adjusted_price ~ carat, data=di)
linear_coefficients <- tidy(linear_fit)
linear_coefficients
```

@@

```{r spline-fit}
spline_fit <- lm(adjusted_price ~ rcs(carat), data=di)
tidy(spline_fit)
```

@@

```{r spline-plot}
carat_sequence <- data.frame(
  carat = seq(min(di$carat), max(di$carat), length=100))
carat_sequence$prediction <- 
  predict(spline_fit, newdata=carat_sequence)
scatter1 + geom_line(
  data=carat_sequence,
  aes(carat, prediction))
```

@@

```{rlinear-plus-spline}
linear_plus_spline_fit <- lm(
  adjusted_price ~ carat + rcs(carat), 
  data=di)
anova(linear_fit, linear_plus_spline_fit)
```

>>>>

```{r diamond-spline-01, echo=FALSE}
library(broom)
library(glue)
library(pspline)
library(readr)
library(rms)
library(tidyverse)
library(yaml)

@@

```{r diamond-spline-02}
f2 <- "singapore-diamond-prices.txt"
raw_data <- read.fwf(
  file=glue("{f0}/{f1}/{f2}"), 
  widths=c(5,2,5,4,5))
di <- raw_data
names(di) <- c(
  "carat",
  "colour",
  "clarity",
  "certification",
  "price"
)
```

@@

```{r diamond-spline-03}
exchange_rate <- 1.72
inflation_rate <- 1.81
di$adjusted_price <- di$price*inflation_rate/exchange_rate
head(di)
```

@@

```{r diamond-spline-04}
summary(di$carat)
summary(di$price)
summary(di$adjusted_price)
count(di, colour)
count(di, clarity)
count(di, certification)
```

@@

```{r diamond-spline-05}
ggplot(data=di, aes(carat, adjusted_price)) +
  geom_point(pch=1) -> scatter1
scatter1
```

@@

```{r diamond-spline-06}
scatter1+geom_smooth(method="lm")
```

@@

```{r diamond-spline-07}
linear_fit <- lm(adjusted_price ~ carat, data=di)
linear_coefficients <- tidy(linear_fit)
linear_coefficients
```

@@

```{r diamond-spline-08}
spline_fit <- lm(adjusted_price ~ rcs(carat), data=di)
tidy(spline_fit)
```

@@

```{r diamond-spline-09}
carat_sequence <- data.frame(
  carat = seq(min(di$carat), max(di$carat), length=100))
carat_sequence$prediction <- 
  predict(spline_fit, newdata=carat_sequence)
scatter1 + geom_line(
  data=carat_sequence,
  aes(carat, prediction))
```

@@

```{r diamond-spline-10}
linear_plus_spline_fit <- lm(
  adjusted_price ~ carat + rcs(carat), 
  data=di)
anova(linear_fit, linear_plus_spline_fit)
```

### 2024-07-07 text/24/different-end-times.Rmd
Updated 2024 Rmd files

```{r}
suppressMessages(library(survival))
suppressMessages(library(tidyverse))
set.seed(190225)
raw_time1 <- rexp(100, rate=0.004)
raw_time2 <- rexp(100, rate=0.008)
sim <- data.frame(
  g=rep(c("a", "b"), each=100), 
  raw_time=c(raw_time1, raw_time2))
head(sim)
tail(sim)
```

@@

```{r}
sim %>%
  group_by(g) %>%
  summarize(
    n_0_to_190=sum(raw_time <= 190),
    n_191_to_225=sum(raw_time > 190&raw_time<=225),
    n_above_225=sum(raw_time > 225)) -> sim_counts
sim_counts
```

@@

```{r}
censor_time_v1 <- rep(c(190, 225), each=100)
sim %>%
  mutate(t_v1=pmin(raw_time, censor_time_v1)) %>%
  mutate(censor_v1=as.numeric(t_v1 < censor_time_v1)) -> sim_v1
sim_v1 %>%
  count(g, censor_v1)
```

@@

```{r}
cox_v1 <- coxph(Surv(t_v1, censor_v1)~g, data=sim_v1)
cox_v1
```

@@

```{r}
censor_time_v2 <- rep(c(190, 190), each=100)
sim %>%
  mutate(t_v2=pmin(raw_time, censor_time_v2)) %>%
  mutate(censor_v2=as.numeric(t_v2 < censor_time_v2)) -> sim_v2
sim_v2 %>%
  count(g, censor_v2)
```

@@

```{r}
cox_v2 <- coxph(Surv(t_v2, censor_v2)~g, data=sim_v2)
cox_v2
```

>>>>

```{r different-end-times-01}
suppressMessages(library(survival))
suppressMessages(library(tidyverse))
set.seed(190225)
raw_time1 <- rexp(100, rate=0.004)
raw_time2 <- rexp(100, rate=0.008)
sim <- data.frame(
  g=rep(c("a", "b"), each=100), 
  raw_time=c(raw_time1, raw_time2))
head(sim)
tail(sim)
```

@@

```{r different-end-times-02}
sim %>%
  group_by(g) %>%
  summarize(
    n_0_to_190=sum(raw_time <= 190),
    n_191_to_225=sum(raw_time > 190&raw_time<=225),
    n_above_225=sum(raw_time > 225)) -> sim_counts
sim_counts
```

@@

```{r different-end-times-04}
censor_time_v1 <- rep(c(190, 225), each=100)
sim %>%
  mutate(t_v1=pmin(raw_time, censor_time_v1)) %>%
  mutate(censor_v1=as.numeric(t_v1 < censor_time_v1)) -> sim_v1
sim_v1 %>%
  count(g, censor_v1)
```

@@

```{r different-end-times-05}
cox_v1 <- coxph(Surv(t_v1, censor_v1)~g, data=sim_v1)
cox_v1
```

@@

```{r different-end-times-06}
censor_time_v2 <- rep(c(190, 190), each=100)
sim %>%
  mutate(t_v2=pmin(raw_time, censor_time_v2)) %>%
  mutate(censor_v2=as.numeric(t_v2 < censor_time_v2)) -> sim_v2
sim_v2 %>%
  count(g, censor_v2)
```

@@

```{r different-end-times-07}
cox_v2 <- coxph(Surv(t_v2, censor_v2)~g, data=sim_v2)
cox_v2
```

### 2024-07-07 text/24/encoding-example.Rmd
Updated 2024 Rmd files

```{r}
#| echo: true
sample_string_1 <- "\x66\x61\x63\x69\x6C\x65"
sample_string_1
```

@@

```{r}
#| echo: true
sample_string_2 <- "\x66\x61\xE7\x69\x6C\x65"
sample_string_2
```

@@

```{r}
#| echo: true
Encoding(sample_string_2)
```

@@

```{r}
#| echo: true
Encoding(sample_string_2) <- "latin1"
sample_string_2
```

@@

```{r}
#| echo:true
sample_string_3 <- 
  iconv(
    sample_string_2, "latin1", "UTF-8")
sample_string_3
```

```{r}
#| echo: true
charToRaw(sample_string_1)
charToRaw(sample_string_2)
charToRaw(sample_string_3)
```

@@

```{r}
#| echo: false
Encoding(sample_string_3) <- "latin1"
sample_string_3
```

>>>>

```{r encoding-example-01}
#| echo: true
sample_string_1 <- "\x66\x61\x63\x69\x6C\x65"
sample_string_1
```

@@

```{r encoding-example-02}
#| echo: true
sample_string_2 <- "\x66\x61\xE7\x69\x6C\x65"
sample_string_2
```

@@

```{r encoding-example-03}
#| echo: true
Encoding(sample_string_2)
```

@@

```{r encoding-example-04}
#| echo: true
Encoding(sample_string_2) <- "latin1"
sample_string_2
```

@@

```{r encoding-example-05}
#| echo:true
sample_string_3 <- 
  iconv(
    sample_string_2, "latin1", "UTF-8")
sample_string_3
```

```{r encoding-example-06}
#| echo: true
charToRaw(sample_string_1)
charToRaw(sample_string_2)
charToRaw(sample_string_3)
```

@@

```{r encoding-example-07}
#| echo: false
Encoding(sample_string_3) <- "latin1"
sample_string_3
```

### 2024-07-07 text/24/nice-layout.Rmd
Updated 2024 Rmd files

---
title: Nice layout for comparison of categorical outcomes
source: "New"
author: Steve Simon
date: 2024-06-27
categories:
- Blog post
tags:
- Being updated
output: html_document
page_update: complete
---

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(glue)
library(mice)
library(readr)
library(rms)
library(tidyverse)
library(yaml)
g0 <- "https://raw.githubusercontent.com/pmean/"
r0 <- "datasets/master/"
f1 <- "titanic.yaml"
s1 <- "yaml-functions.R"
source(paste0(g0, r0, s1))
dd <- read_yaml(paste0(g0, r0, f1))
vnames <- extract_names(dd)
vlabels <- extract_labels(dd)
```

@@

```{r, echo=FALSE}
f2 <- "titanic.txt"
raw_data <- 
  data.frame(
    read_tsv(
      file=paste0(g0, r0, f2),
      col_types="ccncn"))
# Re-arranging columns just for convenient display
o <-  c(1, 3, 2, 4, 5)
ti0 <- raw_data[ , o]
vnames <- vnames[o]
head(ti0)
```

@@

```{r}
ti0 %>%
  group_by(Sex, Survived) %>%
  summarize(n1=n(), .groups="drop") -> tab01
tab01
```

@@

```{r}
tab01 %>%
  group_by(Sex) %>%
  mutate(n2=sum(n1)) %>%
  mutate(pct=round(100*n1/n2)) -> tab02
tab02
```

@@

```{r}
tab02 %>%
  mutate(out=glue("{n1}/{n2} ({pct}%)")) %>%
  select(Sex, Survived, out) -> tab03
tab03
```

@@

```{r}
tab03 %>%
  pivot_wider(
    names_from=Survived, 
    values_from=out) -> tab05
tab05
```

@@

```{r}
tab05 %>%
  set_names(c(
    "Sex", 
    "Deaths", 
    "Survivors")) -> tab06
tab06
```

@@

```{r}
library(kableExtra)
tab06 %>%
  kbl %>%
  kable_paper(
    "striped", 
    full_width=FALSE) -> tab07
tab07
```

@@

```{r}
library(kableExtra)
tab06 %>%
  set_names(c(" ", "Deaths", "Survivors")) %>%
  kbl %>%
  kable_paper(
    "striped", 
    full_width=FALSE) %>%
  pack_rows("Sex", 1, 2) -> tab08
tab08
```

@@

```{r}
ti0 %>%
  group_by(Sex, Survived) %>%
  summarize(n1=n(), .groups = "drop") %>%
  group_by(Sex) %>%
  mutate(n2=sum(n1)) %>%
  mutate(pct=round(100*n1/n2)) %>%
  mutate(out=glue("{n1}/{n2} ({pct}%)")) %>%
  select(Sex, Survived, out) %>%
  pivot_wider(names_from=Survived, values_from=out) %>%
  set_names(c(" ", "Deaths", "Survivors")) %>%
  kbl %>%
  kable_paper("striped", full_width=FALSE) %>%
  pack_rows("Sex", 1, 2)
```

@@

```{r}
ti0 %>%
  group_by(Sex, Survived) %>%
  summarize(n1=n(), .groups = "drop") %>%
  group_by(Sex) %>%
  mutate(n2=sum(n1)) %>%
  mutate(pct=100*n1/n2) %>%
  mutate(out=sprintf("%3.0f/%3.0f (%3.0f%%)", n1, n2, pct)) %>%
  select(Sex, Survived, out) %>%
  pivot_wider(names_from=Survived, values_from=out) %>%
  set_names(c(" ", "Deaths", "Survivors")) %>%
  kbl %>%
  kable_paper("striped", full_width=FALSE) %>%
  pack_rows("Sex", 1, 2)
```

>>>>

---
title: Nice layout for comparison of categorical outcomes
source: "New"
author: Steve Simon
date: 2024-06-27
categories:
- Blog post
tags:
- Incomplete pages
- R programming
- Rmarkdown code
output: html_document
page_update: complete
---

@@

```{r nice-layout-01, echo=FALSE, message=FALSE, warning=FALSE}
library(glue)
library(mice)
library(readr)
library(rms)
library(tidyverse)
library(yaml)
g0 <- "https://raw.githubusercontent.com/pmean/"
r0 <- "datasets/master/"
f1 <- "titanic.yaml"
s1 <- "yaml-functions.R"
source(paste0(g0, r0, s1))
dd <- read_yaml(paste0(g0, r0, f1))
vnames <- extract_names(dd)
vlabels <- extract_labels(dd)
```

@@

```{r nice-layout-02, echo=FALSE}
f2 <- "titanic.txt"
raw_data <- 
  data.frame(
    read_tsv(
      file=paste0(g0, r0, f2),
      col_types="ccncn"))
# Re-arranging columns just for convenient display
o <-  c(1, 3, 2, 4, 5)
ti0 <- raw_data[ , o]
vnames <- vnames[o]
head(ti0)
```

@@

```{r nice-layout-03}
ti0 %>%
  group_by(Sex, Survived) %>%
  summarize(n1=n(), .groups="drop") -> tab01
tab01
```

@@

```{r nice-layout-04}
tab01 %>%
  group_by(Sex) %>%
  mutate(n2=sum(n1)) %>%
  mutate(pct=round(100*n1/n2)) -> tab02
tab02
```

@@

```{r nice-layout-05}
tab02 %>%
  mutate(out=glue("{n1}/{n2} ({pct}%)")) %>%
  select(Sex, Survived, out) -> tab03
tab03
```

@@

```{r nice-layout-06}
tab03 %>%
  pivot_wider(
    names_from=Survived, 
    values_from=out) -> tab05
tab05
```

@@

```{r nice-layout-07}
tab05 %>%
  set_names(c(
    "Sex", 
    "Deaths", 
    "Survivors")) -> tab06
tab06
```

@@

```{r nice-layout-08}
library(kableExtra)
tab06 %>%
  kbl %>%
  kable_paper(
    "striped", 
    full_width=FALSE) -> tab07
tab07
```

@@

```{r nice-layout-09}
library(kableExtra)
tab06 %>%
  set_names(c(" ", "Deaths", "Survivors")) %>%
  kbl %>%
  kable_paper(
    "striped", 
    full_width=FALSE) %>%
  pack_rows("Sex", 1, 2) -> tab08
tab08
```

@@

```{r nice-layout-10}
ti0 %>%
  group_by(Sex, Survived) %>%
  summarize(n1=n(), .groups = "drop") %>%
  group_by(Sex) %>%
  mutate(n2=sum(n1)) %>%
  mutate(pct=round(100*n1/n2)) %>%
  mutate(out=glue("{n1}/{n2} ({pct}%)")) %>%
  select(Sex, Survived, out) %>%
  pivot_wider(names_from=Survived, values_from=out) %>%
  set_names(c(" ", "Deaths", "Survivors")) %>%
  kbl %>%
  kable_paper("striped", full_width=FALSE) %>%
  pack_rows("Sex", 1, 2)
```

@@

```{r nice-layout-11}
ti0 %>%
  group_by(Sex, Survived) %>%
  summarize(n1=n(), .groups = "drop") %>%
  group_by(Sex) %>%
  mutate(n2=sum(n1)) %>%
  mutate(pct=100*n1/n2) %>%
  mutate(out=sprintf("%3.0f/%3.0f (%3.0f%%)", n1, n2, pct)) %>%
  select(Sex, Survived, out) %>%
  pivot_wider(names_from=Survived, values_from=out) %>%
  set_names(c(" ", "Deaths", "Survivors")) %>%
  kbl %>%
  kable_paper("striped", full_width=FALSE) %>%
  pack_rows("Sex", 1, 2)
```

### 2024-07-07 text/24/survival-lasso.Rmd
Updated 2024 Rmd files

---
title: Illustrating the lasso model using the Worcester Heart Attack Study
source: "New"
author: Steve Simon
date: 2024-05-07
categories:
- Blog post
tags:
- Survival analysis
- R programming
output: html_document
---

>>>>

---
title: Illustrating the lasso model using the Worcester Heart Attack Study
source: "New"
author: Steve Simon
date: 2024-05-07
categories:
- Blog post
tags:
- Survival analysis
- R programming
- Rmarkdown graphs
output: html_document
page_update: complete
---

### 2024-07-07 text/24/survival-spline.Rmd
Updated 2024 Rmd files

---
title: Illustrating splines using the Worcester Heart Attack Study
source: "New"
author: Steve Simon
date: 2024-05-07
categories:
- Blog post
tags:
- Survival analysis
- R programming
output: html_document
---

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(broom)
library(pspline)
library(readr)
library(rms)
library(tidyverse)
library(yaml)
f0 <- "https://raw.githubusercontent.com/pmean/datasets/master/"
f1 <- "whas100.yaml"
dd <- read_yaml(paste0(f0, f1))
```

@@

```{r}
f2 <- "whas100.dat"
raw_data <- read.fwf(
  file=paste0(f0, f2),
  widths=c(7, 12, 12, 5, 9, 5, 7, 4, 8))
wa <- raw_data
names(wa) <- c(
  "id",
  "admitdate",
  "foldate",
  "los",
  "lenfol",
  "fstat",
  "age",
  "gender",
  "bmi")
head(wa)
tail(wa)
wa <- wa[-101, ]
wa$lenfol <- wa$lenfol/365.25
```

@@

```{r}
summary(wa$id)
summary(wa$los)
summary(wa$lenfol)
count(wa, fstat)
summary(wa$age)
count(wa, gender)
summary(wa$bmi)
```

@@

```{r}
plot(Surv(wa$lenfol, wa$fstat))
```

```{r}
linear_fit <- coxph(
  Surv(wa$lenfol, wa$fstat)~age,
  data=wa)
linear_fit
```

```{r}
spline_fit <- coxph(
  Surv(wa$lenfol, wa$fstat)~rcs(age),
  data=wa)
spline_fit
```

```{r}
age_sequence <- data.frame(
  age = seq(min(wa$age), max(wa$age), length=100))
age_sequence$prediction <- 
  predict(spline_fit, newdata=age_sequence, type="lp")
ggplot(data=age_sequence, aes(age, prediction)) +
  geom_line()
```

>>>>

---
title: Illustrating splines using the Worcester Heart Attack Study
source: "New"
author: Steve Simon
date: 2024-05-07
categories:
- Blog post
tags:
- Incomplete pages
- Survival analysis
- R programming
- Rmarkdown graphs
output: html_document
page_update: complete
---

@@

```{r survival-spline-01, echo=FALSE, message=FALSE, warning=FALSE}
library(broom)
library(pspline)
library(readr)
library(rms)
library(tidyverse)
library(yaml)
f0 <- "https://raw.githubusercontent.com/pmean/datasets/master/"
f1 <- "whas100.yaml"
dd <- read_yaml(paste0(f0, f1))
```

@@

```{r survival-spline-02}
f2 <- "whas100.dat"
raw_data <- read.fwf(
  file=paste0(f0, f2),
  widths=c(7, 12, 12, 5, 9, 5, 7, 4, 8))
wa <- raw_data
names(wa) <- c(
  "id",
  "admitdate",
  "foldate",
  "los",
  "lenfol",
  "fstat",
  "age",
  "gender",
  "bmi")
head(wa)
tail(wa)
wa <- wa[-101, ]
wa$lenfol <- wa$lenfol/365.25
```

@@

```{r survival-spline-03}
summary(wa$id)
summary(wa$los)
summary(wa$lenfol)
count(wa, fstat)
summary(wa$age)
count(wa, gender)
summary(wa$bmi)
```

@@

```{r survival-spline-04}
plot(Surv(wa$lenfol, wa$fstat))
```

```{r survival-spline-05}
linear_fit <- coxph(
  Surv(wa$lenfol, wa$fstat)~age,
  data=wa)
linear_fit
```

```{r survival-spline-06}
spline_fit <- coxph(
  Surv(wa$lenfol, wa$fstat)~rcs(age),
  data=wa)
spline_fit
```

```{r survival-spline-07}
age_sequence <- data.frame(
  age = seq(min(wa$age), max(wa$age), length=100))
age_sequence$prediction <- 
  predict(spline_fit, newdata=age_sequence, type="lp")
ggplot(data=age_sequence, aes(age, prediction)) +
  geom_line()
```

### 2024-07-07 text/24/titanic-imputation.Rmd
Updated 2024 Rmd files

---
title: Illustrating imputation of missing data using the Titanic dataset
source: "New"
author: Steve Simon
date: 2024-05-07
categories:
- Blog post
tags:
- Missing data
- R programming
output: html_document
---

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(glue)
library(mice)
library(readr)
library(rms)
library(tidyverse)
library(yaml)
g0 <- "https://raw.githubusercontent.com/pmean/"
r0 <- "datasets/master/"
f1 <- "titanic.yaml"
s1 <- "yaml-functions.R"
source(paste0(g0, r0, s1))
dd <- read_yaml(paste0(g0, r0, f1))
vnames <- extract_names(dd)
vlabels <- extract_labels(dd)
```

@@

```{r, echo=FALSE}
f2 <- "titanic.txt"
raw_data <- 
  data.frame(
    read_tsv(
      file=paste0(g0, r0, f2),
      col_types="ccncn"))
# Re-arranging columns just for convenient display
o <-  c(1, 3, 2, 4, 5)
ti0 <- select(raw_data, o)
vnames <- vnames[o]
head(ti0)
```

@@

```{r, echo=FALSE}
cat(vnames[1], ": ", vlabels[1], "\n")
summary(ti0$Age)
cat(vnames[2], ": ", vlabels[2], "\n")
count(ti0, PClass)
cat(vnames[3], "\n")
count(ti0, Sex)
cat(vnames[4], "\n")
cat(as.yaml(dd$vars$Survived$values), "\n")
count(ti0, Survived)
```

@@

```{r}
missing_rows <- which(is.na(ti0$Age))[1:3]
ti0 %>% slice(missing_rows)
```

@@

```{r}
ti1 <- select(ti0, -Name)
ti1$i_female <- as.numeric(ti1$Sex=="female")
ti1 <- select(ti1, -Sex)
ti_mice <- mice(ti1)
```

@@

```{r}
ti_mice %>%
  complete(1) %>%
  slice(missing_rows)
```

@@

```{r}
ti_mice %>%
  complete(2) %>%
  slice(missing_rows)
ti_mice %>%
  complete(3) %>%
  slice(missing_rows)
ti_mice %>%
  complete(4) %>%
  slice(missing_rows)
ti_mice %>%
  complete(5) %>%
  slice(missing_rows)

@@

```{r}
ti_with <- with(
  ti_mice, 
  glm(
    Survived~PClass+rcs(Age, 3)+i_female, 
    family=binomial)) 
```

@@

```{r}
ti_with$analyses[[1]]
```

@@

```{r}
ti_with$analyses[[2]]
ti_with$analyses[[3]]
ti_with$analyses[[4]]
ti_with$analyses[[5]]
```

@@

```{r}
ti_pool <- pool(ti_with)
ti_summary <- summary(ti_pool)
ti_summary
```

@@

```{r}
female <- ti_summary$term=="i_female"
log_or <- ti_summary[female, "estimate"]
se <- ti_summary[female, "std.error"]
or <- round(exp(log_or), 1)
lo <- round(exp(log_or-1.96*se), 1)
hi <- round(exp(log_or+1.96*se), 1)
glue("Odds ratio for females is {or}, 95% CI ({lo},{hi})")
```

>>>>

---
title: Illustrating imputation of missing data using the Titanic dataset
source: "New"
author: Steve Simon
date: 2024-05-07
categories:
- Blog post
tags:
- Missing data
- R programming
- Rmarkdown code
output: html_document
page_update: complete
---

@@

```{r titanic-imputation-01, echo=FALSE, message=FALSE, warning=FALSE}
library(glue)
library(mice)
library(readr)
library(rms)
library(tidyverse)
library(yaml)
g0 <- "https://raw.githubusercontent.com/pmean/"
r0 <- "datasets/master/"
f1 <- "titanic.yaml"
s1 <- "yaml-functions.R"
source(paste0(g0, r0, s1))
dd <- read_yaml(paste0(g0, r0, f1))
vnames <- extract_names(dd)
vlabels <- extract_labels(dd)
```

@@

```{r titanic-imputation-02, echo=FALSE}
f2 <- "titanic.txt"
raw_data <- 
  data.frame(
    read_tsv(
      file=paste0(g0, r0, f2),
      col_types="ccncn"))
# Re-arranging columns just for convenient display
o <-  c(1, 3, 2, 4, 5)
ti0 <- select(raw_data, o)
vnames <- vnames[o]
head(ti0)
```

@@

```{r titanic-imputation-03, echo=FALSE}
cat(vnames[1], ": ", vlabels[1], "\n")
summary(ti0$Age)
cat(vnames[2], ": ", vlabels[2], "\n")
count(ti0, PClass)
cat(vnames[3], "\n")
count(ti0, Sex)
cat(vnames[4], "\n")
cat(as.yaml(dd$vars$Survived$values), "\n")
count(ti0, Survived)
```

@@

```{r titanic-imputation-04}
missing_rows <- which(is.na(ti0$Age))[1:3]
ti0 %>% slice(missing_rows)
```

@@

```{r titanic-imputation-05}
ti1 <- select(ti0, -Name)
ti1$i_female <- as.numeric(ti1$Sex=="female")
ti1 <- select(ti1, -Sex)
ti_mice <- mice(ti1)
```

@@

```{r titanic-imputation-06}
ti_mice %>%
  complete(1) %>%
  slice(missing_rows)
```

@@

```{r titanic-imputation-07}
ti_mice %>%
  complete(2) %>%
  slice(missing_rows)
ti_mice %>%
  complete(3) %>%
  slice(missing_rows)
ti_mice %>%
  complete(4) %>%
  slice(missing_rows)
ti_mice %>%
  complete(5) %>%
  slice(missing_rows)

@@

```{r titanic-imputation-08}
ti_with <- with(
  ti_mice, 
  glm(
    Survived~PClass+rcs(Age, 3)+i_female, 
    family=binomial)) 
```

@@

```{r titanic-imputation-09}
ti_with$analyses[[1]]
```

@@

```{r titanic-imputation-10}
ti_with$analyses[[2]]
ti_with$analyses[[3]]
ti_with$analyses[[4]]
ti_with$analyses[[5]]
```

@@

```{r titanic-imputation-11}
ti_pool <- pool(ti_with)
ti_summary <- summary(ti_pool)
ti_summary
```

@@

```{r titanic-imputation-12}
female <- ti_summary$term=="i_female"
log_or <- ti_summary[female, "estimate"]
se <- ti_summary[female, "std.error"]
or <- round(exp(log_or), 1)
lo <- round(exp(log_or-1.96*se), 1)
hi <- round(exp(log_or+1.96*se), 1)
glue("Odds ratio for females is {or}, 95% CI ({lo},{hi})")
```
